---
title: "SVF_YoungAged_high_resolution"
output: html_document
date: "2024-09-09"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Subtype analysis for macrophage and T cells in Young-Aged SVF mix test 

This vignette is to show the minimal data processing steps for macrophage and T cell subtype analysis in Young-Aged SVF cell mixing test. This script includes 2 sections: \

1. High resolution cell subtype for macrophages;\ 
2. Pseudotime analysis for T cell subtypes. 

#2. Pseudotime analysis for T cell subtypes 
```{r ,message=FALSE,warning=FALSE} 

#subset T cells 

Idents(SVFya_soup)<-SVFya_soup$seurat_clusters
T_SVF<-subset(SVFya_soup,subset= seurat_clusters== "1"|seurat_clusters== "2"|seurat_clusters== "5"|seurat_clusters== "9"|seurat_clusters== "18")

#remove negatives or multiplets, and here we show the analysis of T cells from aged group

Idents(T_SVF)<-T_SVF$deMULTIplex
T_A_SVF<-subset(T_SVF,idents="HTOAged")

T_Y_SVF<-subset(T_SVF,idents="HTOYoung")

#load Monocle3 library (Trapnell et al.)
library(monocle3)
library(SeuratWrappers)
T_traj<-as.cell_data_set(T_SVF) 

fData(T_traj)$gene_short_name<-rownames(fData(T_traj))
reacreate.partition <- c(rep(1,length(T_traj@colData@rownames)))
names(reacreate.partition) <- T_traj@colData@rownames
reacreate.partition <- as.factor(reacreate.partition)

#Use Monocle3 to analyze T cell trajectory

T_traj@clusters$UMAP$partitions <- reacreate.partition
list_cluster <- T_A_SVF$seurat_clusters
T_traj@clusters$UMAP$clusters <- list_cluster

T_traj@int_colData@listData$reducedDims$UMAP <- T_A_SVF@reductions$umap@cell.embeddings

cluster.before.trajectory <- plot_cells(T_traj,
                                        color_cells_by = 'cluster',
                                        label_groups_by_cluster = FALSE,
                                        group_label_size = 8,
                                       cell_size = 1 ) +
  theme(legend.position = "right")

cluster.names <- plot_cells(T_traj,
                            color_cells_by = "seurat_clusters",
                            label_groups_by_cluster = FALSE,
                            group_label_size = 5,
                            cell_size = 3) +
  scale_color_manual(values = c('red', 'blue', 'green', 'maroon')) +
  theme(legend.position = "right")


T_traj <- learn_graph(T_traj, use_partition = FALSE)
plot_cells(T_traj,
           color_cells_by = 'cluster',
           label_groups_by_cluster = FALSE,
           label_branch_points = FALSE,
           label_roots = FALSE,
           label_leaves = FALSE,
           group_label_size = 5,
           cell_size = 1)

#cluster 18 has high expression of Stmn and Pclaf, which are reported as cycling T cell features (Cottam et al.)
T_traj <- order_cells(T_traj, reduction_method = 'UMAP', root_cells = colnames(T_traj[,clusters(T_traj) == '18']))

plot_cells(T_traj,
           color_cells_by = 'pseudotime',
           label_groups_by_cluster = FALSE,
           label_branch_points = FALSE,
           label_roots = FALSE,
           label_leaves = FALSE,
           cell_size = 1)

plot_cells(T_traj,
           color_cells_by = 'pseudotime',
           label_groups_by_cluster = FALSE,
           label_branch_points = FALSE,
           label_roots = FALSE,
           label_leaves = FALSE,
           cell_size = 1,
           trajectory_graph_color = "grey",
            group_label_size = 5) 

Idents(T_A_SVF)<-T_A_SVF$seurat_clusters
CD8_LvsE <- FindMarkers(T_A_SVF, 
                         ident.1 = c("2","9"), 
                         ident.2 = "5",
                         assay = "SCT", 
                         slot = "data",
                         min.pct = 0.25, 
                         logfc.threshold = 0.1)

CD8_LvsE<- CD8_LvsE %>% mutate(logP = -log10(p_val), signif = ifelse(p_val < 0.01 & abs(avg_log2FC)>1, "significant", "not significant"))
CD8_LvsE$gene<-rownames(CD8_LvsE)

T_genes_to_label <- CD8_LvsE%>% 
  filter(signif=="significant")

p<- ggplot(CD8_LvsE, aes(x = avg_log2FC, y = logP, color = signif)) +
  geom_point(alpha = 0.5) +
  theme_minimal() +
  labs(title = "Volcano plot of DE genes in CD8+ T cells", x = "Log2 Fold Change", y = "-Log10 p-value") +
  geom_text(data = T_genes_to_label, aes(label = gene), vjust = 1.5, hjust = 0.5, check_overlap = TRUE, size = 3) +
  scale_color_manual(values = c("significant" = "red", "not significant" = "black")) 


```

## Including Plots


```{r pressure, echo=FALSE}


```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
